---
title: "Homework 4"
author: "Ethan Thibault"
date: "2/7/2018"
output: html_document
---

1. Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the value of the variable z and print the value stored in z.

```{r}
# assign variables
x <- 1.1
a <- 2.2
b <-3.3

# a
z = x^a^b
z

# b
z = (x^a)^b
z


# c
z = 3 * x^3 + 2 * x^2 + 1
z

# d

# Use floor to get the first digit of the value
c = floor(z)

# Subtract that value to leave you with only the numbers after the decimal point
d = z - c

# Multiply that number by 10 to move the next digit into the ones place
e = d * 10

# Get the first digit with with floor again
z = floor(e)
z
```

2. Using the rep and seq functions, create the following vectors:

(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)

(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)

(5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)

```{r}

# concatenate two vectors, each made using seq
vec1 = c(seq(from=1,to=8),seq(from=7,to=1))
vec1

# use rep to repeat elements in a vector and use times to indicate the number of reps per element
vec2 = rep(x=(seq(from=1,to=5)),times=1:5)
vec2

# use same method as above
vec3 = rep(x=(seq(from=5,to=1)),times=1:5)
vec3
```

3. Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as x and y coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R’s inverse trigonometry functions (asin(), acos(), or atan()), convert these numbers into polar coordinates (If you don’t know what polar coordinates are, read about them on the web or in your calculus textbook).

```{r}

# Generate random uniform numbers
vec1=runif(2)

# use formulas to transform x and y cartesian coordinates into polar coordinates (r and theta)
r = sqrt(vec1[1]^2 + vec1[2]^2)
theta = atan(vec1[2]/vec1[1])
r
theta
```

4. Suppose that queue <- c("sheep", "fox", "owl", "ant") and that queue represents the animals that are lined up to enter Noah’s Ark, with the sheep at the front of the line. Using R expressions, update the queue successively as

- the serpent arrives;
- the sheep enters the ark;
- the donkey arrives and talks his way to the front of the line;
- the serpent gets impatient and leaves;
- the owl gets bored and leaves;
- the aphid arrives and the ant invites him to cut in line.
- finally, determine the position of the aphid in the line.

```{r}

# create initial queue by concatinating strings
queue <- c("sheep", "fox", "owl", "ant")
queue

# add serpent
queue <- c(queue, "serpent")
queue

# remove sheep
queue <- queue[queue!="sheep"]
queue

# add donkey to the beginning of the list by concatination
queue <- c("donkey",queue)
queue

# remove serpent
queue <- queue[queue!="serpent"]
queue

# remove owl
queue <- queue[queue!="owl"]
queue

# insert aphid before ant by removing ant from the end of the vector and concatinating aphid and ant on the end of queue
queue <- c(queue[queue!="ant"], "aphid","ant")
queue

# get position of aphid with which function
posAphid <- which(queue=="aphid")
posAphid
```

5. Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7.

```{r}

# create vector from 1 to 100
vec1 <- seq(from=1,to=100)

# remove everything not divisible by 2 (produces a remainder not equal to zero)
vec2 <- vec1[(vec1%%2)!=0]

# remove everything not divisible by 3 (produces a remainder not equal to zero)
vec3 <- vec2[(vec2%%3)!=0]

# remove everything not divisible by 7 (produces a remainder not equal to zero)
vec4 <- vec3[(vec3%%7)!=0]
vec4
```

6. Create a vector z of 1000 random uniform numbers.
- create a vector that contains 3 numbers: the proportion of the numbers in z that are less than 0.10, greater than 0.90, and between 0.45 and 0.55.
- Making successive copies of z, transform your vector of uniform numbers in the following ways:
  - log (base 10) of z
  - z2
  - ez
  - square root of z
- for each case calculate your vector of 3 numbers to get the new proportions.

```{r}

# generate a vector of 1000 random uniform values
z <- runif(1000)

# put anything less than 0.1 into a vector
less <- z[z<0.1]

# put anything more than 0.9 into a vector
more <-z[z>0.9]

# put anything between 0.45 and 0.55 into a vector
between1 <- z[z>0.45]
betweenf <- between1[between1<0.55]

# get proportion of each by dividing length by total numbers
l <- length(less)/1000
m <- length(more)/1000
b <- length(betweenf)/1000

# contatinate proportions
vec1 <- c(l,m,b)
vec1

# repeat above four times, but alter random number vector first
z <- runif(1000)
z <- log10(z)
less <- z[z<0.1]
more <-z[z>0.9]
between1 <- z[z>0.45]
betweenf <- between1[between1<0.55]
l <- length(less)/1000
m <- length(more)/1000
b <- length(betweenf)/1000
vec2 <- c(l,m,b)
vec2

z <- runif(1000)
z <- z*2
less <- z[z<0.1]
more <-z[z>0.9]
between1 <- z[z>0.45]
betweenf <- between1[between1<0.55]
l <- length(less)/1000
m <- length(more)/1000
b <- length(betweenf)/1000
vec3 <- c(l,m,b)
vec3

z <- runif(1000)
z <- z*exp(1)
less <- z[z<0.1]
more <-z[z>0.9]
between1 <- z[z>0.45]
betweenf <- between1[between1<0.55]
l <- length(less)/1000
m <- length(more)/1000
b <- length(betweenf)/1000
vec4 <- c(l,m,b)
vec4

z <- runif(1000)
z <- sqrt(z)
less <- z[z<0.1]
more <-z[z>0.9]
between1 <- z[z>0.45]
betweenf <- between1[between1<0.55]
l <- length(less)/1000
m <- length(more)/1000
b <- length(betweenf)/1000
vec5 <- c(l,m,b)
vec5

```
